---
title: "10_Exercises_R-For-DataScientist"
author: "Ghilman Al Fatih"
date: "1/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse')
```


## ------------------- 10.5 Exercise -------------------------
# 1. How can you tell if an object is a tibble? (Hint: try printing `mtcars`, which is a regular data frame).
```{r soal_1_code, echo=FALSE}
# use this in console
mtcars #output full table (dataframe)
as_tibble(mtcars) #only output 10 first row
is_tibble(mtcars) #FALSE
```

# 2. Compare and contrast the following operations on a `data.frame` and equivalent `tibble`. What is different? Why might the default data frame behaviours cause you frustration?
frustation on selecting the column
```{r soal_2, echo=TRUE}
# as dataframe
df <- data.frame(abc = 1, xyz = "a")
df$x #a
df[, "xyz"] #a
df[, c("abc", "xyz")] # 1 a

# as tibble
df <- as_tibble(df)
df$x #null
df[, "xyz"] #a
df[, c("abc", "xyz")] # 1 a
```

# 3. If you have the name of a variable stored in an object, e.g. `var <- "mpg"`, how can you extract the reference variable from a tibble?
using `df[['var']]` will output the reference variable
using `df['var']` will output another tibble with column `var`

# 4. Practice referring to non-syntactic names in the following data frame by:
1. Extracting the variable called 1.
2. Plotting a scatterplot of 1 vs 2.
3. Creating a new column called 3 which is 2 divided by 1.
4. Renaming the columns to one, two and three.
```{r soal_4_code, echo=FALSE}
# the tibble
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
  
# 1. Extracting the variable called 1
annoying[["1"]]

# 2. Plotting a scatterplot of 1 vs 2
ggplot(annoying, aes(x = `1`, y = `2`)) + geom_point()

# 3. Creating a new column called 3 which is 2 divided by 1.
annoying <- annoying %>%
  mutate(`3` = `2`/`1`)

# 4. Renaming the columns to one, two and three.
annoying %>%
  rename(one = `1`, two = `2`, three = `3`)
```

# 5. What does `tibble::enframe()` do? When might you use it?
I think it's like a dictionary in python, it might be useful
```{r soal_5_code, echo=TRUE}
enframe(1:3)
enframe(c(a = 5, b = 7))
enframe(list(one = 1, two = 2:3, three = 4:6))
deframe(enframe(1:3))
deframe(tibble(a = 1:3))
deframe(tibble(a = as.list(1:3)))

enframe(list(one = 1, two = 2:3, three = 4:6)) %>%
  deframe()

```

# 6. What option controls how many additional column names are printed at the footer of a tibble?
`print(df, n_extra = 2)` print additional 2 columns at the footer
