---
title: "11_Exercise_R_For_DataScientist"
author: "Ghilman Al Fatih"
date: "1/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse')
```

## ------------------- 11.2.2 Exercise -------------------------
# 1. What function would you use to read a file where fields were separated with `|`?
`read_delim('name file.txt', delim = '|')`

# 2. Apart from `file`, `skip`, and `comment`, what other arguments do `read_csv()` and `read_tsv()` have in common?
`quote`, `escape_backslash`, `escape_double`, `col_names`, `col_types`, `locale`, `na`, `quoted_na`, `trim_ws`, `n_max`, `guess_max`, `progress`, `skip_empty_rows`

# 3. What are the most important arguments to `read_fwf()`?
the most important argument is `col_positions` because its specify column position on text file where you define that as a value

# 4. Sometimes strings in a CSV file contain commas. To prevent them from causing problems they need to be surrounded by a quoting character, like " or '. By convention, `read_csv()` assumes that the quoting character will be ", and if you want to change it you’ll need to use `read_delim()` instead. What arguments do you need to specify to read the following text into a data frame?
```{r soal_4_code, echo=TRUE}
file <- "x,y\n1,'a,b'"
read_csv(file, quote = "'")
```

# 5. Identify what is wrong with each of the following inline CSV files. What happens when you run the code?
```{r soal_5, echo=FALSE}
#Last column dropped, no column name
read_csv("a,b\n1,2,3\n4,5,6") 

#Last column on row one become NA, fourth column on second row dropped cause there are only three column
read_csv("a,b,c\n1,2\n1,2,3,4") 

#NA Value on last column because lack of data, and also the quote is ambiguous, so it's treated as doubles, not char
read_csv("a,b\n\"1")

#because last row is char, so the first row treated as char too even without quotation
read_csv("a,b\n1,2\na,b")

#because this is csv, not csv2, the `;` is not treated as delimiter
read_csv("a;b\n1;3")
```


## ------------------- 11.3.5 Exercise -------------------------
# 1. What are the most important arguments to `locale()`?
`encoding` for text
`decimal_mark` and `grouping_mark` for number
`tz` for time zone
`date_names`, `date_format`, `time_format` for date and time format or language

# 2. What happens if you try and set `decimal_mark` and `grouping_mark` to the same character? What happens to the default value of `grouping_mark` when you set `decimal_mark` to “,”? What happens to the default value of `decimal_mark` when you set the `grouping_mark` to “.”?
```{r soal_2_code, echo=FALSE}
# same decimal and grouping mark, ERROR
# locale(grouping_mark = ".", decimal_mark = ".")

# set decimal mark to "," -> grouping_mark become "." by default
locale(decimal_mark = ",")

# set grouping mark to "," -> decimal mark become "." by default
locale(grouping_mark = ",")
```

# 3. I didn’t discuss the `date_format` and `time_format` options to `locale()`. What do they do? Construct an example that shows when they might be useful.
It's useful to change date or time format to usual format used in your place.
```{r soal_3_code, echo=TRUE}
# indonesian format of date and time
idn_locale <- locale(date_format = "%d/%m/%Y", 
                     time_format = "%H/%M/%S", 
                     date_names(day = c("Minggu (Min)", "Senin (Sen)", "Selasa (Tue)", "Rabu (Rab)", "Kamis (Kam)", "Jumat (Jum)", "Sabtu (Sab)"),
                                mon = c("Januari (Jan)", "Februari (Feb)", "Maret (Mar)", "April (Apr)",
                                        "Mei (Mei)", "Juni (Jun)", "Juli (Jul)", "Agustus (Agu)", "September (Sep)",
                                        "Oktober (Okt)", "November (Nov)", "Desember (Des)")),
                     decimal_mark = ",",
                     grouping_mark = ".",
                     tz = "Asia/Jakarta"
                     )
parse_date("03/02/2016", locale = idn_locale)

#default
parse_date("2016-02-03") #default US
parse_date("03/02/2016") #Error, can't read Indonesian format

```

# 4. If you live outside the US, create a new locale object that encapsulates the settings for the types of file you read most commonly.
I've already created it before, for date, date_name, and time
```{r soal_4_code, echo=TRUE}
idn_locale
```

# 5. What’s the difference between `read_csv()` and `read_csv2()`?
`read_csv()` use comma
`read_csv2` use ";" (semicolon)

# 6. What are the most common encodings used in Europe? What are the most common encodings used in Asia? Do some googling to find out.
most common encoding in the world: UTF-8
most common in Europe: UTF-8, ISO 8859
most common in Asia: UTF-8

# 7. Generate the correct format string to parse each of the following dates and times:
```{r soal_7_code, echo=TRUE}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
t1 <- "1705"
t2 <- "11:15:10.12 PM"

parse_date(d1, "%B %d, %Y")
parse_date(d2, "%Y-%b-%d")
parse_date(d3, "%d-%b-%Y")
parse_date(d4, "%B %d (%Y)")
parse_date(d5, "%m/%d/%y")
parse_time(t1, "%H%M")
parse_time(t2, "%H:%M:%OS %p")
```
