---
title: "15_Exercise_R-For-DataScientist"
author: "Ghilman Al Fatih"
date: "2/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(forcats)
```


# ------------------- 15.3.1 Exercise -------------------------
## 1. Explore the distribution of `rincome` (reported  income). What makes the default bar chart hard to understand? How could you improve the plot?

default chart is hard to understand because of overlapping labels, too much variable on horizontal axis, there are `no answer` response that don't need to be included in the chart. The factor levels also confusing
```{r soal_1_code}
gss_cat %>%
  ggplot(aes(x = rincome)) + 
  geom_bar()

```

I propose this change to improve the readability of the chart.

```{r soal_1_code_2}
gss_cat %>% 
  filter(!rincome %in% c('Not applicable', 'Refused', "Don't know", 'No answer')) %>%
  mutate(rincome = fct_recode(rincome, "Less than $1000" = "Lt $1000"),
         rincome = factor(rincome, levels = rev(levels(rincome)))) %>%
  ggplot(aes(x = rincome)) +
  geom_bar() + 
  coord_flip() +
  labs(y = "Number of Report", x = "Reported Income Range", title = "Distribution of Reported Income")
```

## 2. What is the most common `relig` in this survey? What’s the most common `partyid`?
The most common religion is protestant and common party is independent
```{r soal_2_code}
gss_cat %>%
  count(relig) %>% arrange(desc(n)) %>% head(1)

gss_cat %>%
  count(partyid) %>% arrange(desc(n)) %>% head(1)
```

## 3. Which `relig` does `denom` (denomination) apply to? How can you find out with a table? How can you find out with a visualisation?
First we remove all no responses answer
```{r soal_3_code}
unique(gss_cat$relig) #to remove -> None, No answer, Don't know
unique(gss_cat$denom) #to remove -> No answer, Don't know, No denomination, Not applicable
```

Then we figure it out which religion does denomination apply to. The answer is Protestant
```{r soal_3_code_2}
gss_cat %>%
  filter(!denom %in% c('Not applicable', "Don't know", 'No answer', 'No denomination') & !relig %in% c('None','No answer', "Don't know")) %>%
  distinct(relig)
```

For finding out with visualization. We can count how many denomination that each religion have. But it turns out that x axis only have one religion, which is Protestant.
```{r soal_3_code_3}
gss_cat %>%
  filter(!denom %in% c('Not applicable', "Don't know", 'No answer', 'No denomination') & !relig %in% c('None','No answer', "Don't know")) %>%
  count(relig, denom) %>%
  ggplot(aes(x = relig, y = denom, size = n)) +
  geom_point()
```


# ------------------- 15.4.1 Exercise -------------------------
## 1. There are some suspiciously high numbers in `tvhours`. Is the mean a good summary?
I think the mean is not a good summary, I prefer median. Because mean easily affected by outlier.
```{r soal_1_code}
gss_cat %>%
  ggplot() +
  geom_histogram(aes(x = tvhours), na.rm = TRUE, binwidth = 1)

```

## 2. For each factor in `gss_cat` identify whether the order of the levels is arbitrary or principled.
There are six factors in `gss_cat` which is `marital`, `race`, `rincome`, `partyid`, `relig`, `denom`

`marital` is arbitrary. No order in level, even in frequency of marital status
```{r marital}
levels(gss_cat$marital)
gss_cat %>% ggplot(aes(marital)) + geom_bar()
```

`race` is principled. It ordered by frequency of race
```{r race}
gss_cat %>% ggplot(aes(race)) + geom_bar()
levels(gss_cat$race)
```

`rincome` is principled. It ordered from higher range - less range of income
```{r rincome}
levels(gss_cat$rincome)
```

`partyid` is principled. It ordered from strong republican, independent, and then strong democrat
```{r partyid}
levels(gss_cat$partyid)
```

`relig` is arbitrary. No clear ordering from levels and frequency
```{r relig}
levels(gss_cat$relig)
gss_cat %>% ggplot() + geom_barh(aes(y=relig))

```

`denom` is arbitrary. No clear order
```{r denom}
levels(gss_cat$denom)
gss_cat %>% ggplot() + geom_barh(aes(y=denom))
```


## 3. Why did moving “Not applicable” to the front of the levels move it to the bottom of the plot?
Because the position on the plot are determined by factor level. So, if we moving "Not applicable" to the front of the levels with `fct_relevel()`, it will move "Not applicable" to the bottom of the plot


# ------------------- 15.5.1 Exercise -------------------------
## 1. How have the proportions of people identifying as Democrat, Republican, and Independent changed over time?
Other: No response or other party
democrat, republican, and independent

The visualisation result in decreasing republican supporter from 2004 to 2012.
Other party have pretty stable proportion
```{r soal_1_code}
gss_cat %>%
  mutate(partyid = fct_collapse(partyid,
                                other = c("No answer", "Don't know", "Other party"),
                                democrat = c("Not str democrat", "Strong democrat"),
                                republican = c("Strong republican", "Not str republican"),
                                independent = c("Ind,near rep", "Independent", "Ind,near dem")
                                )
         ) %>%
  count(year, partyid) %>%
  group_by(year) %>%
  mutate(
    prop = n / sum(n)
  ) %>%
  ggplot(aes(x = year, y = prop, color = partyid)) +
  geom_line() + geom_point()
```

## 2. How could you collapse rincome into a small set of categories?
```{r soal_2_code}
gss_cat %>%
  mutate(
    rincome = 
      fct_collapse(rincome,
                   Unknown = c("No answer", "Don't know", "Refused", "Not applicable"),
                   `Less than 5000` = c("$4000 to 4999", "$3000 to 3999", "$1000 to 2999", "Lt $1000" ),
                   `$5000 - 10000` = c("$8000 to 9999", "$7000 to 7999", "$6000 to 6999", "$5000 to 5999")
                   ),
    rincome = factor(rincome, levels = rev(levels(rincome)))
  ) %>%
  ggplot() +
  geom_barh(aes(y = rincome))
```


