---
title: "19_Exercise_R-For-DataScientist_Fatih"
author: "Ghilman Al Fatih"
date: "2/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

# ------------------- 19.2.1 Exercise -------------------------
## 1. Why is TRUE not a parameter to `rescale01()`? What would happen if `x` contained a single missing value, and `na.rm` was `FALSE`?
```{19.2.1 reference1}
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
```

`TRUE` is not a parameter on `rescale01()`, because there are no use if user customized this value to `FALSE`, it can be dangerous, and can result in `NA` result or error.

Below is example if we set the `na.rm` value to `NA` like this chunk. `finite = TRUE` basically also mean `na.rm = TRUE` so we exclude those to demonstration. The result will return all `NA` values.

```{r 19.2.1 answer1}
rescale01 <- function(x, na.rm = TRUE) {
  rng <- range(x, na.rm = na.rm)
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale01(c(NA, 0, 1, 2, 3), na.rm = FALSE)
```

## 2. In the second variant of `rescale01()`, infinite values are left unchanged. Rewrite `rescale01()` so that `-Inf` is mapped to 0, and `Inf` is mapped to 1.
```{r 19.2.1 answer2}
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  x <- (x - rng[1]) / (rng[2] - rng[1])
  x[x == -Inf] <- 0
  x[x == Inf] <- 1
  x
}

rescale01(c(NA, 0, 1, 2, 3, Inf, -Inf))
```

## 3. Practice turning the following code snippets into functions. Think about what each function does. What would you call it? How many arguments does it need? Can you rewrite it to be more expressive or less duplicative?
```{19.2.1 code3}
mean(is.na(x))

x / sum(x, na.rm = TRUE)

sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE)
```

First function is proportion of NA values
```{r 19.2.1 answer3}
prop_na <- function(x) {
  mean(is.na(x))
}

prop_na(c(NA, 0, 1, 2, 3, Inf))
```

Second function is proportion of values with its total sum
```{r 19.2.1 answer3.2}
prop_sum <- function(x) {
  x / sum(x, na.rm = TRUE)
}

prop_sum(c(NA, 0, 1, 2, 3))
```

Third function is standard deviation / mean
```{r 19.2.1 answer3.3}
std_div_mean <- function(x) {
  sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE)
}

std_div_mean(c(NA, 0, 1, 2, 3))
```

## 4. Follow http://nicercode.github.io/intro/writing-functions.html to write your own functions to compute the variance and skew of a numeric vector.
ok
```{r 19.2.4 answer4}
# compute variance function
variance <- function(x) {
    n <- length(x)
    m <- mean(x)
    (1/(n - 1)) * sum((x - m)^2)
}

# compute skewness function
skewness <- function(x) {
    n <- length(x)
    v <- var(x)
    m <- mean(x)
    third.moment <- (1/(n - 2)) * sum((x - m)^3)
    third.moment/(var(x)^(3/2))
}

# Try it with normal distribution array
y <- rnorm(1000, mean = 0, sd = 1)
variance(y)
var(y)
skewness(y)
mean(y)
sd(y)
```

the function run pretty good, I just know that skewness can be computed wow.

## 5. Write `both_na()`, a function that takes two vectors of the same length and returns the number of positions that have an NA in both vectors.
```{r 19.2.1 answer5}
both_na <- function(x, y) {
  sum(is.na(x) & is.na(y))  # count how many index that have NA in both array
}

both_na(c(NA, 0, 1, 2, NA, NA), c(NA, NA, 0, NA, NA, 1))
```

## 6. What do the following functions do? Why are they useful even though they are so short?
```{19.2.1 code6}
is_directory <- function(x) file.info(x)$isdir
is_readable <- function(x) file.access(x, 4) == 0
```
the first function is used to check whether `x` is a directory or not. It's useful if your code want to explore the input or output in a folder to prevent error.

the second function is used to check whether `x` is readable or not (has permission, can be opened (ex: not a binary file, file exist, etc.)). It's useful to prevent error when we want to use a file.

## 7. Read the complete lyrics to “Little Bunny Foo Foo”. There’s a lot of duplication in this song. Extend the initial piping example to recreate the complete song, and use functions to reduce the duplication.
```{r 19.2.1 answer7}
foo_foo <- function(x){
  first_verse <- "Little Rabbit Foo Foo \nriding through the forest, \nscooping up the field mice \nand bopping them on the head."
  second_verse <- "Down came the Good Fairy, and said, \n\"Little Rabbit Foo Foo, \nI don't like your attitude, \nscooping upthe field mice \nand bopping them on the head!\""

  countdown <- c('three', 'two', 'one')
  for(i in seq(1,3,1)) {
    third_verse <- paste0("I'll give you ", countdown[i], " chances,\nAnd if you don't behave, I will turn you into a goon!\"\nAnd the next day...\n")
    cat(first_verse, second_verse, third_verse, sep = '\n\n')
  }
  cat('You\'ve got no chances left, so I\'m going to turn you into a goonie.')
}
foo_foo()
```


# ------------------- 19.3.1 Exercise -------------------------
## 1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.
```{r 19.3.1 code1}
f1 <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}
f2 <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
f3 <- function(x, y) {
  rep(y, length.out = length(x))
}

f1(c('precondition','prematur','comprehensive'), 'pre')
f2(c(1,2,3,4))
f3(c(8,4,3,1), 9)
```

`f1` : check whether the string start with prefix or not, better name: `has_prefix()`
`f2` : drop the latest index of an array, if `length(array) <=1` return `NULL`, better name: `drop_last()`
`f3` : repeat y for each x, better name: `repeats()`

## 2. Take a function that you’ve written recently and spend 5 minutes brainstorming a better name for it and its arguments.
```{r 19.3.1 answer2}
# standardise an array of number by removing the mean and scaling to its variance
standardise <- function(x) {
  (x - mean(x)) / sd(x)
}
standardise(rnorm(100))
```

## 3. Compare and contrast `rnorm()` and `MASS::mvrnorm()`. How could you make them more consistent?
`rnorm()` generate samples from the univariate normal distribution, `mvrnorm` generate samples from the multivariate normal distribution.
Could make it more consistent if we change the name to `rnorm()` and `rmvnorm()`

## 4. Make a case for why `norm_r()`, `norm_d()` etc would be better than `rnorm()`, `dnorm()`. Make a case for the opposite.
`norm_r()` and `norm_d()` is good, if we type 'norm', it will suggest `norm_r()` and `norm_d()` (not like the `rnorm`).  It is good if what we thinking first was 'normal distribution', 
`rnorm()` and `dnorm()` on the other hand is hard for suggestion in typing. But good to come up if the first thing we thought was 'random' or 'density'


# ------------------- 19.4.4 Exercise -------------------------
## 1. What’s the difference between `if` and `ifelse()`? Carefully read the help and construct three examples that illustrate the key differences.
`if` only test single condition, if we want to test each element, use `ifelse()`

## 2. Write a greeting function that says “good morning”, “good afternoon”, or “good evening”, depending on the time of day. (Hint: use a time argument that defaults to `lubridate::now()`. That will make it easier to test your function.)
```{r 19.4.4 answer2}
greeting <- function(time = lubridate::now()) {
  hour <- lubridate::hour(time)
  
  if(hour >= 4 && hour < 11  ) {
    "Good Morning"
  } else if (hour >= 11 && hour < 16) {
    "Good Afternoon"
  } else if (hour >= 16 && hour <= 18) {
    "Good Evening"
  } else {
    "Good Night"
  }
}

greeting()
greeting(lubridate::ymd_hms('20190220 16:00:00'))
```

## 3. Implement a `fizzbuzz` function. It takes a single number as input. If the number is divisible by three, it returns “fizz”. If it’s divisible by five it returns “buzz”. If it’s divisible by three and five, it returns “fizzbuzz”. Otherwise, it returns the number. Make sure you first write working code before you create the function.
```{r 19.4.4 answer3}
fizzbuzz <- function(x) {
  if ((x %% 5 == 0 ) && (x %% 3 == 0)) {
    'fizzbuzz'
  } else if (x %% 5 == 0) {
    'buzz'
  } else if (x %% 3 == 0) {
    'fizz'
  } else {
    x
  }
}

fizzbuzz(15)
fizzbuzz(9)
fizzbuzz(20)
fizzbuzz(7)
```

## 4. How could you use `cut() `to simplify this set of nested if-else statements?
```{19.4.4 code4}
if (temp <= 0) {
  "freezing"
} else if (temp <= 10) {
  "cold"
} else if (temp <= 20) {
  "cool"
} else if (temp <= 30) {
  "warm"
} else {
  "hot"
}
```

How would you change the call to cut() if I’d used < instead of <=? What is the other chief advantage of cut() for this problem? (Hint: what happens if you have many values in temp?)

if we use `<=`
```{r 19.4.4 answer4}
(temp <- seq(-20, 60, by = 5))
cut(temp, c(-Inf, 0, 10, 20, 30, Inf),
  right = TRUE,
  labels = c("freezing", "cold", "cool", "warm", "hot")
)
```

if we use `<`
```{r 19.4.4 answer4.2}
(temp <- seq(-20, 60, by = 5))
cut(temp, c(-Inf, 0, 10, 20, 30, Inf),
  right = FALSE,
  labels = c("freezing", "cold", "cool", "warm", "hot")
)
```

## 5. What happens if you use `switch()` with numeric values?
if the input is integer, then the output is case of those number.
if the input is float, then the output is case of `floor(input)`
```{r 19.4.4 answer5}
switch(1, 1, 2, 3)
switch(2, 1, 2, 3)

switch(1.9, 1, 2, 3)
switch(2.6, 1, 2, 3)

```


## 6. What does this `switch()` call do? What happens if `x` is “e”?
```{r 19.4.4 code6}
function6 <- function(x) { switch(x, 
  a = ,
  b = "ab",
  c = ,
  d = "cd"
  )
}
```

Experiment, then carefully read the documentation.

So, the function will return 'ab' if the input is 'a' or 'b', and will return 'cd' if the input is 'c' and 'd'. will return `NULL` if the input is 'e' 'z' or anything else than 'a','b','c','d'
```{r 19.4.4 answer6}
function6('a')
function6('b')
function6('c')
function6('d')
function6('e')
function6('z')
```


# ------------------- 19.5.1 Exercise -------------------------
## 1. What does `commas(letters, collapse = "-")` do? Why?
It will not work. I changed the function a little bit. What that `commas()` do is concat a set of input separated by comma. But I added argument on `collapse` so user can customize the separator
```{r 19.5.1 answer1}
commas <- function(..., collapse = ", ") stringr::str_c(..., collapse = collapse)
commas(letters, collapse = '-')
```

## 2. It’d be nice if you could supply multiple characters to the pad argument, e.g. `rule("Title", pad = "-+")`. Why doesn’t this currently work? How could you fix it?
```{r 19.5.1 reference}
rule <- function(..., pad = "-") {
  title <- paste0(...)
  width <- getOption("width") - nchar(title) - 5
  cat(title, " ", stringr::str_dup(pad, width), "\n", sep = "")
}
rule("Important Rules", pad = '-+')
```

so the output will messed up because of different length of character. so the duplication width is messed up. Here I edited the function so the duplication width will consider the pad length too

```{r 19.5.1 answer2}
rule <- function(..., pad = "-") {
  title <- paste0(...)
  width <- getOption("width") - nchar(title) - 5
  pad_length <- stringr::str_length(pad)
  cat(title, " ", stringr::str_dup(pad, ceiling(width / pad_length)), "\n", sep = "")
}
rule("Important Rules", pad = '-+')
```

## 3. What does the trim argument to `mean()` do? When might you use it?
Basicly what `trim` do is trim the x % first and last sorted array from mean calculation. It can come in handy for calculating mean on skewed distribution.

More information: [link](https://garstats.wordpress.com/2017/11/28/trimmed-means/)

```{r 19.5.1 answer3}
x <- c(39, 92, 75, 61, 45, 87, 59, 51, 87, 12, 8, 93, 74, 16, 32, 39, 87, 12, 47, 50)
trim_coef <- 0.2
xm <- mean(x)
xt <- mean(sort(x)[(1 + floor(trim_coef * length(x))) : (length(x) - floor(trim_coef * length(x)))])
c(xm, mean(x, trim = trim_coef), xt)
```

## 4. The default value for the `method` argument to `cor()` is `c("pearson", "kendall", "spearman")`. What does that mean? What value is used by default?
I think the default value can use three of them. Because "pearson" stated first, it will be used by default. But if "pearson" can't be used on the data, then "kendall", and then "spearman"


